More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-12-21 2016-Weak-FH-Bisim/WeakBisim.tex
changed notation in proof of thm 5 too

        \subsection{Proof of Theorem ~\ref{thm-ctxt-eq}}
 \TODO{Ludo: switch the proof to new notations : s,t instead of s1 s2, and Pred s,t}
The proof of Theorem~\ref{thm-ctxt-eq} exhibits  a bisimulation relation for a 
composed system. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
of $P[Q]_{j_0}$ and obtain an open transition of $P$ on which the FH-bisimulation 
property can 
be applied  to obtain an equivalent family of open transitions of $P'$; this family is 
then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
$P'[Q]_{j_0}$ 
that will simulate the original one.

        Let $\Leaves(Q)=p_l^{l\in L_Q}$, 
$\Leaves(P)=p_l^{l\in L}$, $\Leaves(P')={p'}_l^{l\in L'}$.
	Consider $P$ FH-bisimilar to $P'$. It means that there is a relation 
	$\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
	We will consider the relation $\mathcal{R}'=\{(s_1,s_2|\Pred)|s_1=s'_1\uplus s \land 
	s_2=s'_2\uplus s \land s\in \mathcal{S}_Q \land (s_1',s_2'|\Pred)\in\mathcal{R}\}$ 
	where $\mathcal{S}_Q$ is the set of states of the open automaton of $Q$.	We will prove 
	that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
	states: $(\ostate{s_{1 i}^{i \in L\uplus L_Q}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
	{s}_{1 i}^{i \in L_Q}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
%	automata of 
%	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
Consider an 
	open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
%	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
%	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
\\[-2ex]     
	\[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
		%States(I_1)
		\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
	{ \ostate{s_{1 i}^{i \in L\uplus L_Q}}\OTarrow {v} \ostate{{s'}_{1 i}^{i \in 
	L\uplus 
	L_Q}}}\]
Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
	Lemma~\ref{lem-decompose} we have :\\[-2ex]
			\begin{mathpar}
				P\models{\inferrule*[myfraction=\reddottedrule]
				{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
					\{\xrightarrow{b_j}_j\}^{j\in J'}, 
					\Pred\,',  
					\Post'}
				{\ostate{s_{1 i}^{i\in L}} \OTarrow {v}
					\ostate{s_{1 i}'^{\, i\in L}}}
			}\\
			Q\models{\inferrule*[myfraction=\reddottedrule]
				{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
					\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
					\Post''}
				{\ostate{s_{1 i}^{i\in L_Q}} \OTarrow {v_Q}
					\ostate{s_{1 i}'^{\, i\in L_Q}}}
			}\end{mathpar}
			and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,}'
			\land {\Pred\,}'')$, $\Post_{OT}=\Post'\uplus 
			\Post''$ ($\Post''$ is the restriction of $\Post$ over variables of 
			$\Leaves(Q)$). As $P$ is FH-bisimilar to $P'$ and $(\ostate{s_{1 i}^{i \in 
			L}},\ostate{{s}_{2 i}^{i \in L'}}|\Pred)\in\mathcal{R}$ there is a family 
			$OT'_x$ 
			of 	open transitions of the automaton of $P'$ such that\\[-2ex] 
			\begin{mathpar}
			%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
			P'\models\inferrule*[myfraction=\reddottedrule]
			{\{s_{2 i}\xrightarrow{a_{i x}}_i s_{2 i x}\}^{i\in I_{x}},
				\{\xrightarrow{b_{j x}}_j\}^{j\in J'}, 
				\Pred_{OT_x},\Post_{OT_x}}
			{\ostate{s_{2 i}^{i\in L'}} \OTarrow {v_x} \ostate{s_{2 i x}^{i\in L'}}}
			%         \right)
			\end{mathpar}
			and  $\forall x, (\ostate{s_{1 i}^{i\in L}},\ostate{s_{2 i x}^{i\in 
			L'}}|\Pred_{tgt_x})\in 
			\mathcal{R}$; 
			and  \\
			$\Pred \land \Pred'
		 \Rightarrow \bigvee_{x\in X}
			\left( \forall j\in J'. b_j=b_{jx}  \Rightarrow 
			\Pred_{OT_x}
			\land v\!=\!v_x \land  
			\Pred_{tgt_x}\subst{\Post'}\subst{\Post_{OT_x}}\right)$
			%     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x}) \right)$.

			We can now apply Lemma~\ref{lem-compose} on each of the $OT'_x$ together with 
			the transition of $Q$ and obtain a new family $OT_x$ of open transitions (where for 
			$i\in L_Q$, $s_{2 i}=s_{1 i}$ and $s_{2 i x}=s'_{1 i}$, and for $j\in Holes(Q)$, 
			$b_{j x}=b_j$):\\[-2ex]
				\[ P'[Q]_{j_0}   
				\models
				{\inferrule*[myfraction=\reddottedrule]
					{\{s_{2 i}\xrightarrow{a_{i x}}_i {s}_{2 i x}\}^{i\in I_x \uplus 
					(I\cap L_Q)},
						\{\xrightarrow{b_{j x}}_j\}^{j\in J}, \\
						\Pred_{OT_x}\subst{b_{j_0 x}\leftarrow v_Q}\land\Pred\,'',  
						\Post_{OT_x}\uplus \Post''}
					{\ostate{s_{2 i }^{i\in L'\uplus L_Q}} \OTarrow {v_x}
						\ostate{{s}_{2 i x}^{i\in L'\uplus L_Q}}}
				}
				\]
			Observe that $J=(J\setminus\Holes(Q)\cup \{j_0\})\setminus\{j_0\}\cup 
			(J\cap\Holes(Q))$. Now we have to verify the conditions for the 
			FH-bisimulation between $OT$ and $OT_x$.
			 $\forall x, (\ostate{{s'}_{1 i}^{i\in L\uplus L_Q}},\ostate{s_{2 i 
			x}^{i\in L'\uplus L_Q}}|\Pred_{tgt_x})\in 
			\mathcal{R}'$ (by definition of
                        $\mathcal{R}'$) and in four steps we get:

\noindent                        
\begin{small} $\Pred \land \Pred_{OT} \implies
 \Pred\land(\Pred\,'\land \Pred\,'')\subst{b_{j_0}\leftarrow v_Q}\\ % reversed 
 %substitution 
 \implies 
 (\Pred\land\Pred')\subst{b_{j_0}\leftarrow v_Q}\land \Pred\,''\\ % $b_{j_0}$ only 
 %appears 
 %in $\Pred\,'$
 \implies  \hspace{-2ex}
{\displaystyle{\bigvee_{x\in X}\!(\forall j\!\in\! J'. b_j\!=\!b_{jx}  
\Rightarrow \!
\Pred_{OT_x}\subst{b_{j_0}\!\!\leftarrow\!\! v_Q}
\land v\!=\!v_x}} %\\ \hspace*{40mm} %
\land \Pred_{tgt_x}\subst{\Post'}\subst{\Post_{OT_x}}) \land
\Pred''\\ % 
			%FH for P
 \implies  \hspace{-2ex}
 {\displaystyle{\bigvee_{x\in X}\!( \forall j\!\in\! J'. b_j\!=\!b_{jx} \Rightarrow 
\Pred_{OT_x}\subst{b_{j_0}\!\!\leftarrow\!\! v_Q}\!\land\! \Pred\,''%\\ \hspace*{40mm}
\!\land\! v\!=\!v_x \!\land\!  
\Pred_{tgt_x}\subst{\Post'}\subst{\Post_{OT_x}})}} 			
$\end{small}

@@

Explanations: first, we reversed the substitution; then we
restricted the 
substitution as $b_{j_0}$ only appears in $\Pred\,'$; then we
applied the property of FH-bisimulation for $P$ and rearranged the
terms. The obtained formula reaches the goal except for two points:\\[-4.3ex] 
\begin{itemize}
	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J'$ with 
	$J'\!=\!J\!\setminus\! \Holes(Q) \cup \{j_0\}$ but the formula under the quantifier 
	does not depend on 
	$b_{j_0}$ now (thanks to 
	the substitution). Concerning $\Holes(Q)$, adding prerequisite on more variables 
	does not 
	change the validity of the formula (those variables are not used).
	\item We need $\Pred_{tgt_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}\uplus \Post''}$ 
	, i.e.:
	$\Pred_{tgt_x}\subst{Post'\uplus Post''}\subst{\Post_{OT_x}\uplus \Post''}$. We 
	can conclude by observing that	$\Pred_{tgt_x}$ does not use any variable of $Q$ 
	and thus $\subst{Post''}$ has no effect.
\end{itemize}	
This proves the  condition of the FH-simulation, the other direction is 
similar.\qed

>>>>

        \subsection{Proof of Theorem ~\ref{thm-ctxt-eq}}

The proof of Theorem~\ref{thm-ctxt-eq} exhibits  a bisimulation relation for a 
composed system. It then uses  Lemma~\ref{lem-decompose} to decompose the open transition 
of $P[Q]_{j_0}$ and obtain an open transition of $P$ on which the FH-bisimulation 
property can 
be applied  to obtain an equivalent family of open transitions of $P'$; this family is 
then recomposed by Lemma~\ref{lem-compose} to build a set of open transitions of 
$P'[Q]_{j_0}$ 
that will simulate the original one.

        Let $\Leaves(Q)=p_l^{l\in L_Q}$, 
$\Leaves(P)=p_l^{l\in L}$, $\Leaves(P')={p'}_l^{l\in L'}$.
	Consider $P$ FH-bisimilar to $P'$. It means that there is a relation 
	$\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
	We will consider the relation $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
	t=t'\uplus s'' \land s\in \mathcal{S}_Q \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
	where $\mathcal{S}_Q$ is the set of states of the open automaton of $Q$.	We will prove 
	that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
	states: $(\ostate{s_{1 i}^{i \in L\uplus L_Q}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
	{s}_{1 i}^{i \in L_Q}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
%	automata of 
%	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
Consider an 
	open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
%	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
%	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
\\[-2ex]     
	\[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
		%States(I_1)
		\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
	{ \ostate{s_{i}^{i \in L\uplus L_Q}}\OTarrow {v} \ostate{{s'}_{i}^{~i \in 
	L\uplus 
	L_Q}}}\]
Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
	Lemma~\ref{lem-decompose} we have :\\[-2ex]
			\begin{mathpar}
				P\models{\inferrule*[myfraction=\reddottedrule]
				{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
					\{\xrightarrow{b_j}_j\}^{j\in J'}, 
					\Pred\,',  
					\Post'}
				{\ostate{s_{1}^{i\in L}} \OTarrow {v}
					\ostate{s_{i}'^{~ i\in L}}}
			}\\
			Q\models{\inferrule*[myfraction=\reddottedrule]
				{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
					\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
					\Post''}
				{\ostate{s_{i}^{i\in L_Q}} \OTarrow {v_Q}
					\ostate{s_{i}'^{~ i\in L_Q}}}
			}\end{mathpar}
			and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,}'
			\land {\Pred\,}'')$, $\Post_{OT}=\Post'\uplus 
			\Post''$ ($\Post''$ is the restriction of $\Post$ over variables of 
			$\Leaves(Q)$). As $P$ is FH-bisimilar to $P'$ and $(\ostate{s_{i}^{i \in 
			L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
			$OT'_x$ 
			of 	open transitions of the automaton of $P'$ such that\\[-2ex] 
			\begin{mathpar}
			%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
			P'\models\inferrule*[myfraction=\reddottedrule]
			{\{t_{i}\xrightarrow{a_{i x}}_i t_{i x}\}^{i\in I_{x}},
				\{\xrightarrow{b_{j x}}_j\}^{j\in J'}, 
				\Pred_{OT_x},\Post_{OT_x}}
			{\ostate{t_{i}^{i\in L'}} \OTarrow {v_x} \ostate{t_{i x}^{i\in L'}}}
			%         \right)
			\end{mathpar}
			and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
			L'}}|\Pred_{s x})\in 
			\mathcal{R}$; 
			and  \\
			$\Pred_{s,t} \land \Pred\,'
		 \Rightarrow \bigvee_{x\in X}
			\left( \forall j\in J'. b_j=b_{jx}  \Rightarrow 
			\Pred_{OT_x}
			\land v\!=\!v_x \land  
			\Pred_{s x}\subst{\Post'}\subst{\Post_{OT_x}}\right)$
			%     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x}) \right)$.

			We can now apply Lemma~\ref{lem-compose} on each of the $OT'_x$ together with 
			the transition of $Q$ and obtain a new family $OT_x$ of open transitions (where for 
			$i\in L_Q$, $t_{i}=s_{i}$ and $t_{i x}=s'_{i}$, and for $j\in Holes(Q)$, 
			$b_{j x}=b_j$):\\[-2ex]
				\[ P'[Q]_{j_0}   
				\models
				{\inferrule*[myfraction=\reddottedrule]
					{\{t_{i}\xrightarrow{a_{i x}}_i {t}_{i x}\}^{i\in I_x \uplus 
					(I\cap L_Q)},
						\{\xrightarrow{b_{j x}}_j\}^{j\in J}, \\
						\Pred_{OT_x}\subst{b_{j_0 x}\leftarrow v_Q}\land\Pred\,'',  
						\Post_{OT_x}\uplus \Post''}
					{\ostate{t_{ i }^{i\in L'\uplus L_Q}} \OTarrow {v_x}
						\ostate{{t}_{ i x}^{i\in L'\uplus L_Q}}}
				}
				\]
			Observe that $J=(J\setminus\Holes(Q)\cup \{j_0\})\setminus\{j_0\}\cup 
			(J\cap\Holes(Q))$. Now we have to verify the conditions for the 
			FH-bisimulation between $OT$ and $OT_x$.

			 $\forall x, (\ostate{{s'}_{i}^{~i\in L\uplus L_Q}},\ostate{t_{i 
			x}^{~i\in L'\uplus L_Q}}|\Pred_{s x})\in 
			\mathcal{R}'$ (by definition of
                        $\mathcal{R}'$) and in four steps we get:

\noindent                        
\begin{small} $\Pred_{s,t} \land \Pred_{OT} \implies
 \Pred_{s,t}\land(\Pred\,'\land \Pred\,'')\subst{b_{j_0}\leftarrow v_Q}\\ % reversed 
 %substitution 
 \implies 
 (\Pred\land\Pred\,')\subst{b_{j_0}\leftarrow v_Q}\land \Pred\,''\\ % $b_{j_0}$ only 
 %appears 
 %in $\Pred\,'$
 \implies  \hspace{-2ex}
{\displaystyle{\bigvee_{x\in X}\!(\forall j\!\in\! J'. b_j\!=\!b_{jx}  
\Rightarrow \!
\Pred_{OT_x}\subst{b_{j_0}\!\!\leftarrow\!\! v_Q}
\land v\!=\!v_x}} %\\ \hspace*{40mm} %
\land \Pred_{s x}\subst{\Post'}\subst{\Post_{OT_x}}) \land
\Pred\,''\\ % 
			%FH for P
 \implies  \hspace{-2ex}
 {\displaystyle{\bigvee_{x\in X}\!( \forall j\!\in\! J'. b_j\!=\!b_{jx} \Rightarrow 
\Pred_{OT_x}\subst{b_{j_0}\!\!\leftarrow\!\! v_Q}\!\land\! \Pred\,''%\\ \hspace*{40mm}
\!\land\! v\!=\!v_x \!\land\!  
\Pred_{s x}\subst{\Post'}\subst{\Post_{OT_x}})}} 			
$\end{small}

@@

Explanations: first, we reversed the substitution; then we
restricted the 
substitution as $b_{j_0}$ only appears in $\Pred\,'$; then we
applied the property of FH-bisimulation for $P$ and rearranged the
terms. The obtained formula reaches the goal except for two points:\\[-4.3ex] 
\begin{itemize}
	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J'$ with 
	$J'\!=\!J\!\setminus\! \Holes(Q) \cup \{j_0\}$ but the formula under the quantifier 
	does not depend on 
	$b_{j_0}$ now (thanks to 
	the substitution). Concerning $\Holes(Q)$, adding prerequisite on more variables 
	does not 
	change the validity of the formula (those variables are not used).
	\item We need $\Pred_{s x}\subst{\Post_{OT}}\subst{\Post_{OT_x}\uplus \Post''}$ 
	, i.e.:
	$\Pred_{s x}\subst{\Post'\uplus \Post''}\subst{\Post_{OT_x}\uplus \Post''}$. We 
	can conclude by observing that	$\Pred_{s x}$ does not use any variable of $Q$ 
	and thus $\subst{\Post''}$ has no effect.
\end{itemize}	
This proves the  condition of the FH-simulation, the other direction is 
similar.\qed
